% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amstext}
\usepackage{bm} 
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}
\usepackage{siunitx}
\usepackage{mathrsfs}
\usepackage{wrapfig}
\usepackage{semantic}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{rotating}
\usepackage[dvipsnames]{xcolor}
\usepackage{paralist}
\usepackage[colorlinks]{hyperref}
\usepackage[capitalise,nameinlink]{cleveref}
\usepackage{url}
\usepackage[nounderscore]{syntax}
\usepackage{environ}
\usepackage{listings}
\usepackage{etoolbox}
\usepackage{subcaption}

\AtBeginEnvironment{algorithm}{\linespread{1.05}\selectfont}
\AtBeginEnvironment{grammar}{\small}
\AtBeginEnvironment{grammar}{\linespread{0.6}\selectfont}

\input{commands}



% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}

\title{Co-Simulation is Rewriting Logic
\thanks{We are grateful to the Poul Due Jensen Foundation, which has supported the establishment of a new Centre for Digital Twin Technology at Aarhus University.}}

\author{Simon Thrane Hansen\inst{1} \orcidID{0000-0002-3796-4319} 
\and Peter Csaba Olveczky \inst{2} \orcidID{0000-0002-3796-4319}}
%
%\authorrunning{S. Thrane et al.}

\titlerunning{Co-Simulation using Rewriting Logic}

% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
  DIGIT, Department of Electrical and Computer Engineering, Aarhus University, \\
  %\email{\{sth, claudio.gomes, casper.thule\}@ece.au.dk\} 
  \and Department of Computer Science, Oslo University - \textbf{Check this} \\
}

\maketitle              % typeset the header of the contribution
TACAS:\\
\textbf{Deadline: 14th of October}
\\
\textbf{Page limit: 16}\\
\textbf{Notification: 23rd of December}
\\
WRLA:\\
\textbf{Deadline: 28th of December}
\\
\textbf{Page limit: 15 (without references)}\\
\textbf{Notification: 8th of February}

Formal Framework for Synthesizing Co-simulation algorithms
, analyzing and executing co-simulations
\begin{itemize}
  \item Motivation for co-simulation
  \begin{itemize}
    \item Industry 4.0
    \item Development of CPS
    \item The result of co-simulation
    \begin{itemize}
    \item How to test and verify an OA - need for correct by construction algorithm
    \item Design Space Exploration of the system - model checking 
    \item Testing the configurations of the systems - model checking the contracts.
    \item Dynamic observation of the system
    \item Satisfying the contracts
    \item Checking safety properties (possible to reach a bad state). - model checking might be needed because although each SU might have deterministic behavior, we might not know this - we have to consider a range of possible behaviors.
    \item General problem with writing sound SU.
    \item Time 
  \end{itemize}
  \end{itemize}
  \item Background
  \begin{itemize}
    \item Rewriting Logic
    \item Co-simulation
    \begin{itemize}
      \item 
    \end{itemize}
  \end{itemize}
  \item The Maude model
  \begin{itemize}
    \item 
  \end{itemize}
  \item Analyses
  \begin{itemize}
    \item Confluent and deterministic algorithm
    \item Verification of Algorithms
    \item Design Space Exploration
    \item Step negotiation (Non-deterministic behavior)
    \item Contracts
    \item Interaction with real-world SUs
  \end{itemize}
\end{itemize}

Need for formal methods:
\begin{itemize}
  \item Desireable to do model checking as run-time verification during the simulation.
  \item Communication to external objects - continuous Interaction between formal tool.
  \item Step negotiation
  \item Fixed-point iteration (Algebraic loop)
  \item Complex functions - Tarjan, fixed-point iteration.
\end{itemize}

Contribution
\begin{itemize}
  \item Address the problems - obtaining good co-simulation algorithm
  \item Perform co-simulation in Maude
  \item Step negotiation
  \item Fixed-point iteration (Algebraic loop)
  \item Complex functions - Tarjan, fixed-point iteration.
\end{itemize}

Use of Maude:
\begin{itemize}
  \item Addresses the challenges:
  \item Non-distributed components
  \item Supports functions and is very compressive
  \item Supports model-checking - lets the user explore all different initialization of all the simulation units
  \item Model checking allows one to explore that the SU does not reach a bad state
  \item Most importantly, Maude supports communication with external objects so, in each simulation iteration, we can check and verify (model check) the algorithm used to simulate the next system step.
\end{itemize}

In this paper, we present a co-simulation Synthesizing engine and a Framework for performing various analyses in Maude of both the algorithm and the system of interest.
\begin{itemize}
  \item Model checking approach is used to explore the system's instrumentation's design space - check all possible contracts to reach the best simulation.
  This happens by 
  \item Given a state of the co-simulation system, we synthesize all correct orchestration (one of the algorithms might actually be better than another) for the next step of the co-simulation algorithm.
  \item Model checking is used to show that all synthesized algorithms lead to a  
  \item By defining abstract models of the system in Maude, we can perform model checking to explore violations of safety properties - to 
  \item We then show have this framework can be used to run a co-simulation - to let Maude use as the orchestration engine.
\end{itemize}

We illustrate our techniques with some examples.

Related work
\begin{itemize}
  \item Not a lot of formal methods on co-simulation
  \begin{itemize}
    \item All do work on a symbolic version of the simulation units
    \item Generate only one algorithm
  \end{itemize}
\end{itemize}

\section{Preliminaries}
\subsection{Co-simulation}
\subsection{Maude - Rewriting}
\subsubsection{Model checking}

\section{Formal model of co-simulation in Maude}
Abstract model in Maude 
\begin{itemize}
  \item Inputs
  \item Outputs
  \item Local state
  \item Parameters
\end{itemize}

\section{Synthesizing Algorithm}

\section{Model checking Analysis of co-simulation}
  \subsection{Orchestration Algorithms}
  \subsection{Design Space Exploration for instrumentation of the scenario}

\section{Use Maude an Orchestration Engine}

Interaction with real system

\section{Case study}

\section{Related Work}

\section{Concluding remarks}



TODAY:
\begin{itemize}
  \item Define the Algorithm Verification
  \begin{itemize}
    \item Introduce non-deterministic SUs
  \end{itemize}
  \item Perform various analyses
  \begin{itemize}
    \item Design Space Exploration - instrumentation
    \item Checking Safety Properties
    \item Design Space of the system
  \end{itemize}
  \item Make equation into rules.
\end{itemize}

External objects in Aarhus - ask Casper -maybe implement something in Python:



\section{Progress on Thursday day}
\begin{itemize}
  \item Simulation one synthesized algorithm
  \item Generation of all algorithms - multiple
  \item Show that all algorithms are confluent - one final state
  \item Reachability analyses - can we reach any bad state with any algorithm.
  \item Run for one algorithm - small step semantics
  \item Design Space Exploration:
  \begin{itemize}
    \item Instrumentation of the ports - with no loops and that satisfied certain properties - (show casing two different ways)
    \item Finding the different parameters of the simulation units from a discrete set
    \item Combination of the two.
  \end{itemize}
  \item  Big step simulation running multiple iterations
\end{itemize}

\section{Todo:}

\begin{itemize}
  \item Simulate the system with loops
  \item Rules vs. Equations
  \item External objects - long run
  \item Tidying up
  \item Larger example - to show the different things
\end{itemize}

%\input{abstract.tex}

%\input{introduction.tex}

%\input{background.tex}

%\input{relatedwork.tex}

%\input{maudeModel.tex}

%\input{casestudy.tex}

%\input{conclusion.tex}
%
% ---- Bibliography ---


\bibliographystyle{splncs04}
\bibliography{bib}

\end{document}
